---
title: "RiskCom"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RiskCom}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi=240
)
```

```{r setup}
library(RiskCom)
```

## Introduction

RiskCom is a package containing tools to help in communicating risk.

Previous packages were less customisable and required the use of specific icon libraries. RiskCom is flexible and allows images in files or internet URLs to be used, there is also a small built-in library of custom made icons.

## `waffles`: Waffle plots 

Waffle plots are useful for showing percentages, with each icon representing one percent. RiskCom contains some icons which can be used. Images in files or an internet URL can also be used.

### Single

It is very easy to create simple waffle plots. Here, the small built-in RiskCom library is used. To use a local image or an image from the internet, the `icon_lib` option can be set to `file` or `internet`.

```{r, fig.width=7, fig.height=5, out.width="100%"}
icons = c("happy", "neutral", "sad") # Choose icons
legend = c("Happy", "Neutral", "Sad") # Name for the icons
percentages = c(35, 45, 20) # Percentage of each icon
waffles(icons, legend, percentages, icon_lib="RiskCom") # The waffle plot function
```

### Double

A double waffle plot can be used to add an extra dimension to the waffle plots. The example below shows how many people have a disease and how many people are happy.

```{r, fig.width=7, fig.height=5, out.width="100%"}
icons1 = c("person", "person_grey")
icons2 = c("happy", "neutral", "sad")
legend1 = c("No disease", "Disease")
legend2 = c("Happy", "Neutral", "Sad")
percentages = c(70,9,1, 3, 5, 12)
waffles(icons=icons1, legend=legend1, percentages,
        icons2=icons2, legend2=legend2, icon_size=0.09)
```

## `ninN`: Showing n in N patients with a disease or condition

These can help in visually demonstrating statements such as "Two in five people have high cholesterol." They are similar to waffle plots, but on a smaller scale.

```{r, fig.height=1, fig.width=5, out.width="100%"}
ninN(n=2, N=5, icon="person", plt_size=.8)
```

If, for example, some risk factor increased this to three in five, we could show this using the `additional` argument. Here we also show how the `colours` argument can be used to choose the colour of the icons.

```{r, fig.height=1, fig.width=5, out.width="100%"}
ninN(n=2, N=5, icon="person", additional=1, plt_size=.8, colours = c("black", "blue", "lightgrey"))
```

## `barprob`: Visual plot of probabilties

Pie charts are commonly used, but they can be difficult to read as we find it difficult to judge angles and find it easier to judge straight line distances. An alternative to a bar chart is to just use a single line. The `barprob` function can be used for this. 

```{r}
bar_prob(c("Yes", "No"), 
         c(80,20))

```

```{r}
bar_prob(c("Yes", "Maybe", "No"), 
         c(50,30,20),
         colours=c("green", "orange", "red"))
```

## `bartab`: Displaying a bar chart and a table

For some information, some people prefer a bar chart and other prefer a table. This is an easy way to show both. It is less useful for patients.


### Single

```{r, fig.height=7, fig.width=7, out.width="100%"}
data(mtcars) # Load the car dataset
cyl = factor(mtcars$cyl) # Convert cylinders to factors
gears = factor(mtcars$gear) # Convert gears to factors
bartab(vals = cyl, xlab = "Cylinders", ylab = "Frequency") # The bartab function
```

### Multiple

This also works for multiple bar charts. 

```{r, fig.height=7, fig.width=7, out.width="100%"}
bartab(vals = cyl, xlab = "Cylinders", ylab = "Frequency",
       vals2 = gears, flab = "Gears")
```
